Periodograms for Multiband Astronomical Time Series

### 2015-05-12 writeup/paper.tex
typo in abstract

\begin{abstract}
This paper introduces the {\it multiband periodogram}, a general extension of the well-known 
Lomb-Scargle approach for detecting periodic signals in time-domain data. In addition to 
advantages of the Lomb-Scargle method such as treatment of non-uniform sampling and
heteroscedastic errors, the multiband periodogram significantly improves period finding for randomly 
sampled multiband light curves (e.g., Pan-STARRS, DES and LSST). The light curves in 
each band are modeled as arbitrary truncated Fourier series, with the period and phase 
shared across all bands. The key aspect is the use of Tikhonov regularization which drives
most of the variability into the so-called base model common to all bands, while 
fits for individual bands describe residuals relative to the base model and typically require
lower-order Fourier series. This decrease in the effective model complexity is the
main reason for improved performance. \new{After a pedagogical development of the formalism
of least-squares spectral analysis which motivates the essential features of the multiband model,}
we use simulated light curves 
and randomly subsampled SDSS Stripe 82 data to demonstrate the superiority of this 
method compared to other methods from the literature, and find that this method will be able to
efficiently determine the correct period in the majority of LSST's bright RR Lyrae stars with
as little as six months of LSST data, \new{a vast improvement over the 10+ years of data
reported to be required by previous studies.}
A Python implementation of this method, along with code to fully reproduce the results
reported here, is available on GitHub.
\end{abstract}

>>>>

\begin{abstract}
This paper introduces the {\it multiband periodogram}, a general extension of the well-known 
Lomb-Scargle approach for detecting periodic signals in time-domain data. In addition to 
advantages of the Lomb-Scargle method such as treatment of non-uniform sampling and
heteroscedastic errors, the multiband periodogram significantly improves period finding for randomly 
sampled multiband light curves (e.g., Pan-STARRS, DES and LSST). The light curves in 
each band are modeled as arbitrary truncated Fourier series, with the period and phase 
shared across all bands. The key aspect is the use of Tikhonov regularization which drives
most of the variability into the so-called base model common to all bands, while 
fits for individual bands describe residuals relative to the base model and typically require
lower-order Fourier series. This decrease in the effective model complexity is the
main reason for improved performance. \new{After a pedagogical development of the formalism
of least-squares spectral analysis which motivates the essential features of the multiband model,}
we use simulated light curves 
and randomly subsampled SDSS Stripe 82 data to demonstrate the superiority of this 
method compared to other methods from the literature, and find that this method will be able to
efficiently determine the correct period in the majority of LSST's bright RR Lyrae stars with
as little as six months of LSST data, \new{a vast improvement over the years of data
reported to be required by previous studies.}
A Python implementation of this method, along with code to fully reproduce the results
reported here, is available on GitHub.
\end{abstract}
